Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-04-24 bbbq_article.tex
Processed feedback, thanks Frans and Geert, thanks @bogeert and @fransbianchi

\geert{wat text gewist want dit is de methods sectie}
In previous work, it was found that the over-presentation of TMH-derived peptides can be
explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
Here, a similar analysis was applied, by correlating the percentage of predicted TMH-derived epitopes versus the mean hydrophobicity of all 
peptides.

This study differs in one important aspect from our previous work (\cite{bianchi2017}). 
The definition of a binder differs from \cite{bianchi2017}:
in the current study, a peptide is called a binder if, for a certain haplotype, 
any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
all peptides within a \emph{proteome} 
(see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
whereas the previous study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
This revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes.
To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials. \geert{refereer hier naar supplementary figuur}

@@

% \paragraph{Selection of SNPs}
%
SNPs were picked based on ID number, which is linked to their discovery date. To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined. This analysis showed no positional bias of the SNPs, as shown in \geert{dit is supplement toch?} supplementary figure \ref{fig:snp_rel_pos}.

@@

Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
for a human, viral and bacterial proteome.
Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
with at least one residue.
The horizontal line shows the expected percentage of TMH-derived epitopes 
that would be presented, if TMH-derived epitopes would be presented just as 
likely as epitopes derived from soluble regions.
For 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are predicted to be presented more often 
than the null expectation, for a human and bacterial proteome.
For the viral proteome, 12 out of 13 haplotypes present
TMH-derived epitopes more often than expected by chance.
The extent of the over-presentation between the different haplotypes 
is similar for the probed proteomes, 
which strengthens our previous conclusion (\cite{bianchi2017}) 
that the hydrophobicity of the MHC-binding groove 
is the main factor responsible for the predicted over-presentation 
of TMH-derived peptides.  \frans{zoals eerder HLA-C mist en ook HLA-DP voor de nu gekozen haplotypes mis ik de rationale waarom we deze hebben genomen? volgens mij had ik een artikel gestuurd met de meest voorkomende HLA-DP haplotypes. De rationale voor HLA-C is dat de expressie hier naar verhousding heel laag van is en dat dus een minder explecitie rol speelt in antigen presentatie} \geert{in methods section geeft Richel al rationale voor de keuze van de haplotypes, waarin hij schrijft dat dit gebasseerd is op onze eerdere studie (voor MHC-I) en de frequentie van de haplotypes (voor MHC-II). Dit lijkt mij voldoende.}

@@

  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes} \geert{een figuur heeft altijd een titel met de conclusie. De legenda zelf is feitelijke opsomming en biedt geen (of weinig) interpretatie}
    \textbf{(A)} 
    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The red lines indicate the percentages as expected by chance.
    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B-C)}
    Correlation between the percentages of predicted TMH-derived epitopes
    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. Labels are shorthand for the HLA haplotypes,
    see Table \ref{tab:haplotype_abbreviations} for the names. \frans{wat is de reden dat nr 15 en 8 er buiten vallen? de rest is vrijwel linear)
  }}
\end{figure}

@@

For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}. Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. \frans{dit is dus omdat 2 haplotypes buiten de boot vallen} This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
  \caption{ \textbf{MHC-I and -II presentation of TMH-derived peptides in peptide elution studies.}
    TMH-derived epitopes are presented \emph{in vivo},
    for both MHC-I and MHC-II, for two topology prediction
    programs. \geert{legenda is veel te summier. Wat is precies geplot? ook studies van de data citeren.}
  }
  \label{fig:elution}
\end{figure}

To obtain experimental confirmation that peptides stemming from TMHs 
are presented in MHC-I and MHC-II,
two peptide elution studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} for MHC-I and MHC-II respectively) 
were reanalyzed.
Figure \ref{fig:elution} shows the percentages of epitopes derived 
from TMHs
found in the MHC-I and MHC-II elution 
studies, \geert{is dit van totale peptides, of van humane peptides? Uiteraard van humane peptides gebruiken, want de rest heb je niet gemapped. Dat moet dan wel hier gezegd worden. Daarnaast: zijn de MHC-I en -II haplotypes bekend in deze studies?}
for the two topology prediction tools \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
Regardless of the prediction tool, 
at least 100 epitopes were predicted to be derived from a TMH for each condition. 
From these findings, it is robustly predicted that
epitopes derived from TMHs are presented in both MHC-I and MHC-II.
See the supplementary materials (table \ref{tab:elution}) for the exact values.

@@

  \caption{ \textbf{Evolutionary conservation of human TMHs.}
    \textbf{(A)} 
    The number of SNPs in TMHs as expected by chance (left bar) 
    and found in the dbSNP database (right bar).
    Percentages show the relative conservation
    of SNPs in TMHs found.
    \textbf{(B)}
    Percentage of SNPs found in TMHs.
    Each point shows for one protein the predicted percentage of
    TMH ($x$-axis) and the observed occurrence of SNPs being located
    within a TMH ($y$-axis).
    The dashed diagonal line shows the line of equality (i.e.,
    equal conservation of TMHs and soluble protein regions). The red line shows the linear regression analysis and the grey shaded area the 95\% confidence interval. \geert{please correct last sentence}
    \textbf{(C)}
    Distribution of the percentages of TMH in the TMPs used in this study.
  }
\end{figure}

@@

  \begin{subfigure}[t]{\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
    \label{fig:f_snps_found_and_expected_per_spanner}
  \end{subfigure}
  \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
      \textbf{(A)} 
      The number of SNPs in TMHs as expected by chance 
      and observed in the dbSNP database, 
      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
      Percentages show the relative conservation
      of SNPs in TMHs found.
      \textbf{(B)} 
      Distribution of the proportion of amino acids residing
      in the plasma membrane. 
      \textbf{(C)} 
      Percentage of SNPs found in TMPs predicted to have only a single
      (left) or multiple (right) TMHs.
        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis). The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions). The red lines shows the linear regression analyses and the grey shaded areas the 95\% confidence intervals. \geert{please correct last sentence}
     }
\end{figure}

@@

We also determined the probability to find the 
observed amount of SNPs  by chance in single- and multi-spanners.
For single-spanners, we found 452 SNPs in TMH, where
$\approx462$ were expected by chance. 
The chance to observe this or a lower number by chance is 
$0.319$. As this chance was higher than our $\alpha = 0.025$,
we consider this no significant effect.
For the single-spanners, we found 3,351 SNPs in TMH, where \frans{er staat 2keer single spanners dit zijn. toch de multispanners?}
$\approx3,678$ were expected by chance. 
The chance to observe this or a lower number by chance is 
$8.315841 \cdot 10^{-12}$, 
which means this number is significantly less as explained by variation.

@@

A first possibility is that the extraction of TMPs from the membrane is mediated by the ER-associated degradation (ERAD) machinery. For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER). The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) \cite{rock2016present} first assemble and stabilize the heavy and light chains of MHC-I. Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC). The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}. Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC. In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving} \cite{meusser2005erad}. Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading. This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.

A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded. Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}. The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}. However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein. Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well \cite{lautscham2001processing}, suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.

A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}. TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes. Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles. This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles. Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}. Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs \cite{kleijmeer2001reorganization}\cite{peters1991segregation}\cite{zwart2005spatial}. However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs \cite{zwart2005spatial}, indicating that epitope loading of MHC-II also occurs at intraluminal vesicles. This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact. A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.

Alternatively to the enzymatic degradation of lipids in MVBs by lipases\cite{sander2016lipase} \cite{gilleron2016lysosomal}, they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} and might thereby lead to the extraction of TMPs. Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental evidence for MHC presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\geert{Deze gehele subsection kan wat mij betreft wel weg (of naar de 'minor discussion' in de supplement). Het is deels een herhaling van resultaten (1e en 3e paragraaf). DRiPs in paragraaf 2 kan echt niet alles verklaren, want MHC-II is opname van heel antigen. In parafraag 4, zal de methodological bias zal alleen maar resulteren in het onderschatten van presentatie van TMHs, en is dus ook niet echt relevant hier.}
% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
From reanalysis of peptide elution studies, we show that 1.4\% of epitopes that are naturally presented on MHC-I 
can be mapped to stem from a TMH. 
We also find that MHC-II presents epitopes derived from TMHs, 
as we find that 3.9-4.0\% of sequenced epitopes
can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}). \geert{zie eerdere vraag. Zijn deze percentages van totale peptides, of van humane peptides?}

% \paragraph{False positives}

A caveat of the detected TMH peptides in elution studies is that we assume that 
the found epitopes are all derived from their completely 
translated predecessor proteins.
However, defective ribosomal 
products (DRiPs) (\cite{yewdell1996defective}),
in which the ribosome terminates its translation prematurely  (\cite{yewdell2011drips}), may result in peptide fragments that did not originate from the mature TMPs.
DRiPs may lead to false positives, as we incorrectly deduce
epitopes to be derived from a TMH, instead of from incompletely formed proteins.

Another caveat is that the TMH prediction tools are imperfect.
This imperfection, however, appears to be of limited importance,
as our data show that two different topology prediction software algorithms give similar results, 
as shown by table \ref{tab:elution}).

A final caveat in the elution studies is that 
there is methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less efficiently 
by mass spectronomy (\cite{bianchi2017}). 
This means that the number of TMH-derived epitopes
might be underestimated.

>>>>

In previous work, it was found that the over-presentation of TMH-derived peptides can be
explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
Here, a similar analysis was applied, by correlating the percentage of predicted TMH-derived epitopes versus the mean hydrophobicity of all 
peptides.

This study differs in one important aspect from our previous work (\cite{bianchi2017}). 
The definition of a binder differs from \cite{bianchi2017}:
in the current study, a peptide is called a binder if, for a certain haplotype, 
any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
all peptides within a \emph{proteome} 
(see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
whereas the previous study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
This revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes.
To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials.

@@

% \paragraph{Selection of SNPs}
%
SNPs were picked based on ID number, which is linked to their discovery date. To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined. This analysis showed no positional bias of the SNPs, as shown in supplementary figure \ref{fig:snp_rel_pos}.

@@

Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
for a human, viral and bacterial proteome.
Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
with at least one residue.
The horizontal line shows the expected percentage of TMH-derived epitopes 
that would be presented, if TMH-derived epitopes would be presented just as 
likely as epitopes derived from soluble regions.
For 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are predicted to be presented more often 
than the null expectation, for a human and bacterial proteome.
For the viral proteome, 12 out of 13 haplotypes present
TMH-derived epitopes more often than expected by chance.
The extent of the over-presentation between the different haplotypes 
is similar for the probed proteomes, 
which strengthens our previous conclusion (\cite{bianchi2017}) 
that the hydrophobicity of the MHC-binding groove 
is the main factor responsible for the predicted over-presentation 
of TMH-derived peptides.

@@

  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
    \textbf{(A)} 
    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The red lines indicate the percentages as expected by chance.
    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B-C)}
    Correlation between the percentages of predicted TMH-derived epitopes
    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. Labels are shorthand for the HLA haplotypes,
    see Table \ref{tab:haplotype_abbreviations} for the names.
  }
\end{figure}

@@

For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}. Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
  \caption{
    \textbf{
      Robust prediction that TMH epitopes are presented \emph{in vivo}.
    }
    Bars show the percentage of peptides obtained from elution
    studies that is derived from a TMH, for MHC-I and -II,
    for two TMH prediction tools.
  }
  \label{fig:elution}
\end{figure}

To obtain experimental confirmation that peptides stemming from TMHs 
are presented in MHC-I and MHC-II,
two peptide elution studies were reanalyzed:
For MHC-I, peptides presented \emph{in vivo} 
by the (humane) haplotypes HLA-A and B were sequenced (\cite{schellens2015comprehensive}),
for MHC-II these were haplotypes DQ2.5, DQ2.2, and 
DQ7.5 (\cite{bergseng2015different}).
Figure \ref{fig:elution} shows the percentages of epitopes derived 
from TMHs
found in the MHC-I and MHC-II elution 
studies,
for the two topology prediction tools \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
Regardless of the prediction tool, 
at least 100 epitopes were predicted to be derived from a TMH for each condition. 
From these findings, it is robustly predicted that
epitopes derived from TMHs are presented in both MHC-I and MHC-II.
See the supplementary materials (table \ref{tab:elution}) for the exact values.

@@

  \caption{ \textbf{Evolutionary conservation of human TMHs.}
    \textbf{(A)} 
    The number of SNPs in TMHs as expected by chance (left bar) 
    and found in the dbSNP database (right bar).
    Percentages show the relative conservation
    of SNPs in TMHs found.
    \textbf{(B)}
    Percentage of SNPs found in TMHs.
    Each point shows for one protein the predicted percentage of
    TMH ($x$-axis) and the observed occurrence of SNPs being located
    within a TMH ($y$-axis).
    The dashed diagonal line shows the line of equality (i.e.,
    equal conservation of TMHs and soluble protein regions). The red line indicates a linear fit, the gray area its 95\% confidence interval.
    \textbf{(C)}
    Distribution of the percentages of TMH in the TMPs used in this study.
  }
\end{figure}

@@

  \begin{subfigure}[t]{\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
    \label{fig:f_snps_found_and_expected_per_spanner}
  \end{subfigure}
  \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
      \textbf{(A)} 
      The number of SNPs in TMHs as expected by chance 
      and observed in the dbSNP database, 
      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
      Percentages show the relative conservation
      of SNPs in TMHs found.
      \textbf{(B)} 
      Distribution of the proportion of amino acids residing
      in the plasma membrane. 
      \textbf{(C)} 
      Percentage of SNPs found in TMPs predicted to have only a single
      (left) or multiple (right) TMHs.
        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis). The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions). The red line indicates a linear fit, the gray area its 95\% confidence interval.
     }
\end{figure}

@@

We also determined the probability to find the 
observed amount of SNPs  by chance in single- and multi-spanners.
For single-spanners, we found 452 SNPs in TMH, where
$\approx462$ were expected by chance. 
The chance to observe this or a lower number by chance is 
$0.319$. As this chance was higher than our $\alpha = 0.025$,
we consider this no significant effect.
For the multi-spanners, we found 3,351 SNPs in TMH, where 
$\approx3,678$ were expected by chance. 
The chance to observe this or a lower number by chance is 
$8.315841 \cdot 10^{-12}$, 
which means this number is significantly less as explained by variation.

@@

A first possibility is that the extraction of TMPs from the membrane is mediated by the ER-associated degradation (ERAD) machinery. For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER). The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) (\cite{rock2016present}) first assemble and stabilize the heavy and light chains of MHC-I. Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC). The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I (\cite{blees2017structure}). Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC. In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation (\cite{preston2017evolving,meusser2005erad}). Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading. This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) (\cite{bougneres2009role}), since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo (\cite{fujimoto2006proteasomal}).

A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded. Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E (\cite{oliveira2015alternative}). The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER (\cite{oliveira2015alternative}). However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein. Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well (\cite{lautscham2001processing}), suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.

A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) (\cite{oliveira2015alternative}). TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes. Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles. This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles. Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded (\cite{gruenberg2020life}). Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs (\cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}). However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs (\cite{zwart2005spatial}), indicating that epitope loading of MHC-II also occurs at intraluminal vesicles. This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact. A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading (\cite{zwart2005spatial}).

Alternatively to the enzymatic degradation of lipids in MVBs by lipases (\cite{sander2016lipase,gilleron2016lysosomal}), they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 (\cite{dingjan2016lipid}). This oxidation can result in a destabilization and disruption of membranes (\cite{dingjan2016lipid}) and might thereby lead to the extraction of TMPs. Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading.
